Online Laterality Battery Pre-Processing - short version

### 2020-02-12 battery-preprocessing_v3_DB.Rmd
Modified to correct computation of p.corr for RDT. Also added z-test for difference in proportions for laterality measures based on accuracy

In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.

@@

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for side
    temp_dat$accurate[myrow] <- sum(tmp2$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp2$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong side, so % correct is meaningful

@@

The data is visualised using the pirate plot function and t.test conducted.

```{r plotdata_ttest, warning=FALSE}
#Same function can be used for various tasks - we specify relevant data/task in the function call by assigning origdat etc to relevant values
dopirate <- function(origdat,task,measure){
# measure for each side
# plot the number of correct responses in each side
pirateplot(formula = X1 ~ side,
           data = origdat,
           main = measure,
           theme = 0,
           pal = "southpark", # southpark color palette
           bean.f.o = .0, # Bean fill
           point.o = .3, # Points
           inf.f.o = .7, # Inference fill
           inf.b.o = .8, # Inference border
           avg.line.o = 1, # Average line
           bar.f.o = .5, # Bar
           point.pch = 21,
           point.cex = .7,
           inf.method= "ci")
# plot L & R to get slope
q <- ggplot(data=origdat, aes(x=side, y=X1, group=subject, color=subject)) +
    geom_line() +
    geom_point() +
    ggtitle(paste0(task,": ",measure," by side"))
q + theme_bw() + theme(legend.position = "none")
#  t test
print(paste0(task,measure))
myt1<-t.test(origdat$X1~ origdat$side, paired = TRUE, alternative = "two.sided")
print(myt1)

@@

@@

makeLI <- function(demogs,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
LIdf <- origdat[keeps]
LIdf <- spread(LIdf, side, X1)
# calculate each participants' lat index
LIdf <- 
  LIdf %>% 
  group_by(subject) %>%
  mutate(thisLI= ((Right - Left)/(Right + Left))*100)

@@

# now create a data frame for lat index 
LIdf <- select(LIdf, "subject", "LI")
# now merge with demographics
demogs <- merge(demogs, LIdf, by= "subject")
lastcol<-length(colnames(demogs))
colnames(demogs)[lastcol]<-mycolname

@@

```{r domakeLI}
#We add Chimeric faces LIs to demographics for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_CF_RT_LI"
polarity <-1

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-1
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI)
plot(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,main='Chimeric LI accuracy')
abline(v=0)
abline(h=0)
text(-10,15,paste0('r = ',round(cor1,3)))
cor2<-cor(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI)
plot(demographics$Day1_CF_RT_LI,demographics$Day2_CF_RT_LI,main='Chimeric LI RT')
abline(v=0)
abline(h=0)
text(-5,15,paste0('r = ',round(cor2,3)))

```
As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. The RT advantage is not as expected. This may be due to the buttons being clicked on screen and the additional time penalty of moving the mouse.

@@

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, rhyme, acuracy, RT (RT only for correct answers).

@@

w<-which(RDT_df1$p.corr<60)
if (length(w)>0)
{RDT_df1$accurate[w]<-NA
RDT_df1$RT[w]<-NA
}
w<-which(RDT_df2$p.corr<60)
if (length(w)>0)
{RDT_df2$accurate[w]<-NA
RDT_df2$RT[w]<-NA
  }
```

@@

#now create LI
origdat$X1 <- origdat$p.corr
mycolname<-"RDT_p.corr_LI"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"RDT_RT_LI"
polarity <- (-1)
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(demographics$RDT_acc_LI))
if(length(w)>0)
{demographics$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero
t.test(demographics$RDT_p.corr_LI)
t.test(demographics$RDT_RT_LI)
#rename column to identify day
myncol <- length(colnames(demographics))
colnames(demographics)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(demographics)[(myncol-1):myncol])

}
cor1<-cor(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,use='complete.obs')
plot(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(demographics$Day1.RDT_RT_LI,demographics$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor2,3)))
```
### T-test as alternative to LI for RDT

This is an exploratory analysis.

@@

#compare t-value with LI : use t = 1.65 as cutoff for sig lateralised
 plot(demographics$tlat_RDT_1,demographics$Day1.RDT_RT_LI)
 abline(v=1.65)
  plot(demographics$tlat_RDT_2,demographics$Day2.RDT_RT_LI)
   abline(v=1.65)
#compare t-value from day 1 and day 2 
     myr <- cor(demographics$tlat_RDT_1,demographics$tlat_RDT_2,use='complete.obs')
    plot(demographics$tlat_RDT_1,demographics$tlat_RDT_2)
   abline(v=1.65) 
   abline(h=1.65)
   text(0,2.5,paste0('r = ',round(myr,3)))
```

@@

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$easy1
mycolname<-"Day1_finger.easy_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$hard1
mycolname<-"Day1_finger.hard_LI"
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$easy2
mycolname<-"Day2_finger.easy_LI"
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$hard2
mycolname<-"Day2_finger.hard_LI"
demographics <- makeLI(demographics,origdat,task,mycolname,polarity,mycutoff)

@@

>>>>

In this section, the correct sequences are marked as 1, incorrect answers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.

@@

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for side
    temp_dat$accurate[myrow] <- sum(tmp2$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong side, so % correct is meaningful
    #need to use tmp1, as this contains errors

@@

Generic function for pirate plot and group-level t-test:
```{r plotdata_ttest, warning=FALSE}
#Same function can be used for various tasks - we specify relevant data/task in the function call by assigning origdat etc to relevant values
dopirate <- function(origdat,task,measure){
# measure for each side
# plot the number of correct responses in each side
pirateplot(formula = X1 ~ side,
           data = origdat,
           main = measure,
           theme = 0,
           pal = "southpark", # southpark color palette
           bean.f.o = .0, # Bean fill
           point.o = .3, # Points
           inf.f.o = .7, # Inference fill
           inf.b.o = .8, # Inference border
           avg.line.o = 1, # Average line
           bar.f.o = .5, # Bar
           point.pch = 21,
           point.cex = .7,
           inf.method= "ci")
# plot L & R to get slope
q <- ggplot(data=origdat, aes(x=side, y=X1, group=subject, color=subject)) +
    geom_line() +
    geom_point() +
    ggtitle(paste0(task,": ",measure," by side"))
q + theme_bw() + theme(legend.position = "none")
#  t test
print(paste0(task,measure))
myt1<-t.test(origdat$X1~ origdat$side, paired = TRUE, alternative = "two.sided")
print(myt1)

@@

The data is visualised using the pirate plot function and t.test for laterality at group level conducted.

@@

makeLI <- function(demogs,origdat,task,mycolname1,mycolname2,polarity,mycutoff){

# mycolname1 and 2 are names for newly created columns for LI and statistic (z)
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
LIdf <- origdat[keeps]
LIdf <- spread(LIdf, side, X1)
# calculate each participants' lat index
LIdf <- 
  LIdf %>% 
  group_by(subject) %>%
  mutate(thisLI= ((Right - Left)/(Right + Left))*100)

@@

#statistic is z score - see http://www.sthda.com/english/wiki/one-proportion-z-test-in-r
LIdf$statistic <-NA
for (i in 1:nrow(LIdf)){
    pN <- (LIdf$Left[i]+LIdf$Right[i])
  p0 <- LIdf$Left[i]/ pN
  q <- 1-p0
  pe <- .5
  LIdf$statistic[i]<-(p0-pe)/sqrt(p0*q/pN)
}

# now create a data frame for lat index 
LIdf <- select(LIdf, "subject", "LI","statistic")
# now merge with demographics
demogs <- merge(demogs, LIdf, by= "subject")
lastcol<-length(colnames(demogs))
colnames(demogs)[(lastcol-1):lastcol]<- c(mycolname1,mycolname2)

@@

```{r domakeLI}
#We add Chimeric faces LIs to demographics for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname1<-"Day1_CF_acc_LI"
mycolname2 <-"Day1_CF_z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname1<-"Day2_CF_acc_LI"
mycolname2 <-"Day2_CF_z"
polarity<- (-1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)

cor1 <- cor(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,use="complete.obs")
plot(demographics$Day1_CF_acc_LI,demographics$Day2_CF_acc_LI,main='Chimeric LI accuracy')
abline(v=0)
abline(h=0)
text(-10,15,paste0('r = ',round(cor1,3)))
cor2<-cor(demographics$Day1_CF_z,demographics$Day2_CF_z,use="complete.obs")
plot(demographics$Day1_CF_z,demographics$Day2_CF_z,main='Chimeric LI.z')
abline(v=0)
abline(h=0)
abline(v=1.65,lty=2,col='red')
abline(v=-1.65,lty=2,col='red')
abline(h=1.65,lty=2,col='red')
abline(h=-1.65,lty=2,col='red')
text(-1,3,paste0('r = ',round(cor2,3)))

cor3 <- cor(demographics$Day1_CF_acc_LI,demographics$Day1_CF_z,use="complete.obs")
plot(demographics$Day1_CF_acc_LI,demographics$Day1_CF_z,main='Chimeric LI vs z time1')
abline(v=0)
abline(h=0)

```
As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. 
Agreement between the z-score and conventional LI is very strong.
Test-retest seems good.

@@

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, rhyme, accuracy, RT (RT only for correct answers).

@@

w<-which(RDT_df1$p.corr<60)
if (length(w)>0)
{RDT_df1$accurate[w]<-NA
RDT_df1$RT[w]<-NA
RDT_df1$p.corr[w]<-NA
}
w<-which(RDT_df2$p.corr<60)
if (length(w)>0)
{RDT_df2$accurate[w]<-NA
RDT_df2$RT[w]<-NA
RDT_df2$p.corr[w]<-NA
  }
```

@@

#now create LI
origdat$X1 <- origdat$p.corr
mycolname1<-"RDT_p.corr_LI"
mycolname2 <- "RDT_pcorr.z"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75

demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(demographics$RDT_acc_LI))
if(length(w)>0)
{demographics$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero in sample as a whole
t.test(demographics$RDT_p.corr_LI)
t.test(demographics$RDT_pcorr.z)
#rename column to identify day
myncol <- length(colnames(demographics))
colnames(demographics)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(demographics)[(myncol-1):myncol])

}
cor1<-cor(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,use='complete.obs')
plot(demographics$Day1.RDT_p.corr_LI,demographics$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))

```
### T-test as alternative to LI for RDT

Advantage of using a t-statistics to represent LI is that we can also identify people whose L-R difference is significantly different from zero.

@@

#compare t-value from day 1 and day 2 
     myr <- cor(demographics$tlat_RDT_1,demographics$tlat_RDT_2,use='complete.obs')
    plot(demographics$tlat_RDT_1,demographics$tlat_RDT_2,xlim=c(-4,4),ylim=c(-4,4))
  abline(v=0)
  abline(h=0)
   abline(v=-1.65,lty=2,col='red') 
   abline(h=-1.65,lty=2,col='red')
    abline(v=1.65,lty=2,col='red') 
   abline(h=1.65,lty=2,col='red')
   text(0,2.5,paste0('r = ',round(myr,3)))
```

@@

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$easy1
mycolname1<-"Day1_finger.easy_LI"
mycolname2<-"Day1_finger.easy_z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)

origdat$X1 <- origdat$hard1
mycolname1<-"Day1_finger.hard_LI"
mycolname2<-"Day1_finger.hard_z"
demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)

origdat$X1 <- origdat$easy2
mycolname1<-"Day2_finger.easy_LI"
mycolname2<-"Day2_finger.easy_z"
demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)

origdat$X1 <- origdat$hard2
mycolname1<-"Day2_finger.hard_LI"
mycolname2<-"Day2_finger.hard_z"
demographics <- makeLI(demographics,origdat,task,mycolname1,mycolname2,polarity,mycutoff)

@@

cor3 <- cor(demographics$Day1_finger.easy_z,demographics$Day2_finger.easy_z,use='complete.obs')
plot(demographics$Day1_finger.easy_z,demographics$Day2_finger.easy_z,main='Finger easy z')
abline(v=0)
abline(h=0)
text(-5,20,paste0('r = ',round(cor1,3)))
